Verification of OpenSSL's SHA3 \keccak\ function using the Software Analysis Workbench

### 2021-11-22 paper/paper.tex
Move imp to figs and move figs to proof section
Restructure sections and file names

\documentclass[runningheads]{llncs}
%
\usepackage{paralist}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{color}
\definecolor{blue}{RGB}{42,0,255}
\definecolor{green}{RGB}{63,127,95}
\definecolor{purple}{RGB}{127,0,85}
\definecolor{lime}{RGB}{114, 148, 0}
\usepackage{listings}
\lstdefinelanguage{Cryptol}{
  keywordstyle=[1]\color{blue}\bfseries,
  keywordstyle=[2]\color{purple}\bfseries,
  keywordstyle=[3]\color{black}\bfseries,
  basicstyle={\color{lime}\scriptsize},
  literate=*
  {*}{{{\color{black}$*$}}}{1}
  {=}{{{\color{black}=}}}{1}
  {+}{{{\color{black}+}}}{1}
  {-}{{{\color{black}$-$}}}{1}
  {\%}{{{\color{black}\%}}}{1}
  {(}{{{\color{black}(}}}{1}
  {)}{{{\color{black})}}}{1}
  {\{}{{{\color{black}\{}}}{1}
  {\}}{{{\color{black}\}}}}{1}
  {[}{{{\color{black}[}}}{1}
  {]}{{{\color{black}]}}}{1}
  {|}{{{\color{black}$|$}}}{1}
  {>}{{{\color{black}$>$}}}{1}
  {<}{{{\color{black}$<$}}}{1}
  {\#}{{{\color{black}\#}}}{1}
  {@}{{{\color{black}@}}}{1}
  {.}{{{\color{black}.}}}{1}
  {\'}{{{\color{black}'}}}{1}
  {:}{{{\color{black}:}}}{1}
  {^}{{{\color{black}$\char`\^$}}}{1}
  {&}{{{\color{black}\textbackslash}}}{1},
  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
  morekeywords=[2]{if, then, else, where},
  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
    vals, loop, state, while, take}
}

@@

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
  Parker Hanson\inst{1} \and
  Benjamin Winters\inst{1} \and
  Eric Mercer\inst{1}\orcidID{0000--0002-2264-2958} \and
  Bret Decker\inst{1}
}
%
\authorrunning{P. Hanson et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
  Brigham Young University, Provo UT 84602, USA
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{abstract}
\keywords{verification \and cryptography \and symbolic execution \and SHA3.}
\end{abstract}

@@

%%\input{preliminaries}

\section{Proof Overview and Results}\label{sec:proof}
\input{proof}

\section{SHA3 FIPS Specification in Cryptol}\label{sec:fips}
\input{fips}

@@

\section{Results and Discussion}\label{sec:results}
\input{results}

>>>>

\documentclass[runningheads]{llncs}
%
\usepackage{paralist}
\usepackage{comment}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{color}
\definecolor{blue}{RGB}{42,0,255}
\definecolor{green}{RGB}{63,127,95}
\definecolor{purple}{RGB}{127,0,85}
\definecolor{lime}{RGB}{114, 148, 0}
\usepackage{listings}
\lstdefinelanguage{Cryptol}{
  keywordstyle=[1]\color{blue}\bfseries,
  keywordstyle=[2]\color{purple}\bfseries,
  keywordstyle=[3]\color{black}\bfseries,
  basicstyle={\color{lime}\scriptsize},
  literate=*
  {*}{{{\color{black}$*$}}}{1}
  {=}{{{\color{black}=}}}{1}
  {+}{{{\color{black}+}}}{1}
  {-}{{{\color{black}$-$}}}{1}
  {\%}{{{\color{black}\%}}}{1}
  {(}{{{\color{black}(}}}{1}
  {)}{{{\color{black})}}}{1}
  {\{}{{{\color{black}\{}}}{1}
  {\}}{{{\color{black}\}}}}{1}
  {[}{{{\color{black}[}}}{1}
  {]}{{{\color{black}]}}}{1}
  {|}{{{\color{black}$|$}}}{1}
  {>}{{{\color{black}$>$}}}{1}
  {<}{{{\color{black}$<$}}}{1}
  {\#}{{{\color{black}\#}}}{1}
  {@}{{{\color{black}@}}}{1}
  {.}{{{\color{black}.}}}{1}
  {\'}{{{\color{black}'}}}{1}
  {:}{{{\color{black}:}}}{1}
  {^}{{{\color{black}$\char`\^$}}}{1}
  {&}{{{\color{black}\textbackslash}}}{1},
  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
  morekeywords=[2]{if, then, else, where},
  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
    vals, loop, state, while, take}
}

@@

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}
\newcommand\keccak{\textsc{Keccak}}

\begin{document}
%
\title{Verification of OpenSSL's SHA3 \keccak\ function using the Software Analysis Workbench}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
  Parker Hanson\inst{1} \and
  Benjamin Winters\inst{1} \and
  Eric Mercer\inst{1}\orcidID{0000-0002-2264-2958} \and
  Bret Decker\inst{1}
}
%
\authorrunning{P. Hanson et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
  Brigham Young University, Provo UT 84602, USA
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{abstract}
\keywords{verification \and cryptography \and symbolic execution \and SHA3.}
\end{abstract}

@@

\section{SHA-3 Overview}
\input{sha-3}

\section{Proof Outline and Results}\label{sec:proof}
\input{proof}

\section{FIPS 202 Specification in Cryptol}\label{sec:fips}
\input{fips202}

@@

### 2021-11-22 paper/proof.tex
Move imp to figs and move figs to proof section
Restructure sections and file names

SHA 3 Details
\begin{compactitem}
  \item Process 1600 bits at a time: $5 * 5 * 64$
  \item State vector is $5*5*64*$
  \item Incoming message padded out to whole number of 1600 bit bit-strings
  \item Processed in order by \texttt{KECCAK}
  \item Keccak turns each block into a state vector (rearranges bit layout in memory)
  \item Keccak has 24 rounds for state vector. 
  \item There are 5 hash functions sequentially applied in each round to iteratively update the state vector
  \item After the last round, keccak turns the state vector back into the bit-string./\item The next unprocessed block is \emph{sponged} (e.g., absorbed) into the current bit-string from the latest keccak and the process is repeated 
  \item The max digest (hash) size is 512 bits in the standard
  \item The final bit-string from keccak after sponging and processing every block in the original message is truncated to the desired digest size 
\end{compactitem}

@@

>>>>

SHA 3 Details
\begin{figure*}[ht]
  \centering
  \includegraphics[width=\linewidth]{figs/proof.png}

  \caption{SHA3 Keccak Proof Structure}
  \label{fig:proofStructure}

\end{figure*}
\begin{figure*}[ht]
  \centering
  \includegraphics[width=\linewidth]{figs/proof2.png}

  \caption{Inner Function Contracts}
  \label{fig:proofStructure2}

\end{figure*}

\begin{compactitem}
  \item Process 1600 bits at a time: $5 * 5 * 64$
  \item State vector is $5*5*64*$
  \item Incoming message padded out to whole number of 1600 bit bit-strings
  \item Processed in order by \texttt{KECCAK}
  \item Keccak turns each block into a state vector (rearranges bit layout in memory)
  \item Keccak has 24 rounds for state vector. 
  \item There are 5 hash functions sequentially applied in each round to iteratively update the state vector
  \item After the last round, keccak turns the state vector back into the bit-string./\item The next unprocessed block is \emph{sponged} (e.g., absorbed) into the current bit-string from the latest keccak and the process is repeated 
  \item The max digest (hash) size is 512 bits in the standard
  \item The final bit-string from keccak after sponging and processing every block in the original message is truncated to the desired digest size 
\end{compactitem}

@@

Talk about the cost of the verification. The amount of effort required to write the specification. Make clear the limits of the result proof.

\begin{compactitem}
\item Visual inspection and limited test vectors for (1) and (2)
\item Only provide proof certificates for message sizes of X to Y for digest sizes of \textbf{FIXME} for (3)
\item Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
\end{compactitem}

Effort (see above and then consider below):
\begin{compactitem}
 \item How long were we working on the Cryptol code?
 \item How long were we working on the SAW proofs?
 \item How long did it take to understand the OpenSSL code?
\end{compactitem}
We'll want to determine how long it took us and then try to infer how long that would have taken had we been working full time (40 hours/week).

Because of the modular design of the SHA-3 specification, the Cryptol implementation and OpenSSL implementation are programmatically more similar. 
The only strong discrepancy is that the OpenSSL implementation organizes the bits of the sponge construction. 
The affect of modular cryptographic design can be seen by the proof completing successfully in the Yices, Z3, and ABC SMT-solvers. 
Previous work on SHA2-256 was only successful with Yices \cite{nfm-us}. Full results are show in table \ref{finaltable}.

\begin{table*}[b]
\caption{Proof Runtimes for SHA-3 with SMT-solvers}\label{finaltable}
\setlength{\tabcolsep}{13.5pt}
\begin{tabular}{|l|r|r|r|}
\hline
\textbf{Function} & \textbf{Yices} & \textbf{Z3} & \textbf{ABC} \\
\hline
Pi          &   0.895 secs &    11.072 secs &    0.417 secs \\
Rho         &   4.509 secs &    15.776 secs &    5.879 secs \\
Theta       &  25.942 secs &    11.345 secs &   14.869 secs \\
Chi         &   5.059 secs &     3.930 secs &    1.327 secs \\
Iota        &  11.048 secs &     8.801 secs &  612.497 secs \\
KeccakF1600 & 745.206 secs & 15187.671 secs & 2910.549 secs \\
\hline
\end{tabular}
\end{table*}
